\projectname: A systematic search for transits in \KT\ data

### 2015-01-27 ms.tex
moving math to appendix

The photometry from the \KT\ extension of NASA's \kepler\ mission is wrought
with systematic effects due to the relative imprecision of the telescope
pointing.
We present a method for searching these light curves by simultaneously
fitting for these systematics and the transit signals of interest.
This method is more computationally expensive than standard search algorithms
but we demonstrate that it can be efficiently implemented and used to
discover transit signals in the existing dataset.
We apply this method to the full Campaign 1 dataset and report a list of XXX
planet candidates transiting YYY stars.

@@

One key realization that is also exploited by the official \kepler\ pipeline
is that the systematic trends due to pointing shifts and other instrumental
effects are shared---with different signs and weights---by all the stars on
the detector.
To exploit this fact, the \project{PDC} component of the \kepler\ pipeline
removes any trends from the light curves that can be fit using a linear
combination of a small number of ``eigen-light curves'' (ELCs) found by
running Principal Component Analysis (PCA) on a set of light curves
\citep{map-pdc1, map-pdc2}.
Similarly, we ran PCA (as implemented by the \project{scikit-learn} project
\citealt{sklearn}) on the full set of Campaign 1 light curves to determine a
basis of representative ELCs.
We chose to keep the top 200 components---many more than are normally
used---for our model described in the following section.
\dfmtodo{do we want to plot some of these light curves here?}

@@

Physically, the motivation for our model---and the PDC model---is that every
star on the detector should be affected by same set of systematic effects.
These are due to things like pointing jitter, temperature variations, and
other sources of PSF modulation and each one will be imprinted in the light
curves of many stars with varying amplitudes and signs.
Therefore, while it is hard to write down a physical generative model for the
systematics, building a data-driven model might be possible.
This intuition is also exploited by other methods that model the systematics
using only empirical centroids \citep{vanderberg-a, crossfield} but out more
flexible model should capture the systematic effects more robustly but there
is a risk of over-fitting.

To avoid over-fitting in our pipeline, we simultaneously fit for the transit
signal and the trends using a rigid model for the signal and a relatively
flexible model for the noise.
Specifically, we model the light curve as being generated by linear
combination of the 200 basis light curves and a ``box'' transit model at a
given period and phase.
Formally, this can be written for the light curve of the $k$-th star as
\begin{eqnarray}\eqlabel{linear-model}
\bvec{f}_k &=& \bvec{A}\,\bvec{w}_k + \mathrm{noise}
\end{eqnarray}
where
\begin{eqnarray}
\bvec{f}_k &=& \left (\begin{array}{cccc}
    f_{k,1} & f_{k,2} & \cdots & f_{k,N}
\end{array}\right )^\T
\end{eqnarray}
is the list aperture fluxes for star $k$ observed at $N$ times
\begin{eqnarray}
\bvec{t} &=& \left (\begin{array}{cccc}
    t_{1} & t_{2} & \cdots & t_{N}
\end{array}\right )^\T \quad.
\end{eqnarray}
In \eq{linear-model}, the design matrix is given by
\begin{eqnarray}
\bvec{A} &=& \left (\begin{array}{cccccc}
    x_{1,1} & x_{2,1} & \cdots & x_{J,1} & 1 & m_\bvec{\theta}(t_1) \\
    x_{1,2} & x_{2,2} & \cdots & x_{J,2} & 1 & m_\bvec{\theta}(t_2) \\
    && \vdots &&&\\
    x_{1,N} & x_{2,N} & \cdots & x_{J,N} & 1 & m_\bvec{\theta}(t_N)
\end{array}\right )
\end{eqnarray}
where the $x_{j,n}$ are the basis ELCs---with the index $j$ running over
components and the index $n$ running over time---and $m_\bvec{\theta}(t)$ is
the transit model
\begin{eqnarray}
m_\bvec{\theta}(t) &=& \left\{\begin{array}{cl}
-1 & \mathrm{if\,}t\,\mathrm{in\,transit} \\
0 & \mathrm{otherwise}
\end{array}\right.
\end{eqnarray}
parameterized by a period, phase, and transit duration (these parameters are
denoted by \bvec{\theta}).

@@

>>>>

The photometry from the \KT\ extension of NASA's \kepler\ mission is
afflicted by systematic effects caused by the relative imprecision of the
telescope pointing, and possibly other spacecraft effects.
We present a method for searching these light curves for evidence of
exoplanets by simultaneously fitting for these systematics and the
transit signals of interest.
This method is more computationally expensive than standard search algorithms
but we demonstrate that it can be efficiently implemented and used to
discover transit signals in the existing dataset.
We apply this method to the full Campaign 1 dataset and report a list of XXX
planet candidates transiting YYY stars.
We also present a list of probably astrophysical false positives, and
the results of some follow-up spectroscopic data on a few of the
planet candidates
The most interesting systems are JJJ and HHH.

@@

One key realization that is also exploited by the official \kepler\ pipeline
is that the systematic trends caused by pointing shifts and other instrumental
effects are shared---with different signs and weights---by all the stars on
the detector.
To exploit this fact, the \project{PDC} component of the \kepler\ pipeline
removes any trends from the light curves that can be fit using a linear
combination of a small number of ``eigen-light curves'' (ELCs) found by
running Principal Component Analysis (PCA) on a set of light curves
\citep{map-pdc1, map-pdc2}.
Similarly, we ran PCA (as implemented by the \project{scikit-learn} project
\citealt{sklearn}) on the full set of Campaign 1 light curves to determine a
basis of representative ELCs.
We chose to keep the top 200 components---many more than are normally
used---for our model described in the following section.
\dfmtodo{do we want to plot some of these light curves here?}

@@

Physically, the motivation for our model---and the PDC model---is that every
star on the detector should be affected by same set of systematic effects.
These are caused by things like pointing jitter, temperature variations, and
other sources of PSF modulation and each one will be imprinted in the light
curves of many stars with varying amplitudes and signs.
Therefore, while it is hard to write down a physical generative model for the
systematics, building a data-driven model might be possible.
This intuition is also exploited by other methods that model the systematics
using only empirical centroids \citep{vanderberg-a, crossfield} but out more
flexible model should capture the systematic effects more robustly but there
is a risk of over-fitting.

To avoid over-fitting in our pipeline, we simultaneously fit for the transit
signal and the trends using a rigid model for the signal and a relatively
flexible model for the noise.
Specifically, we model the light curve as being generated by linear
combination of the 200 basis light curves and a ``box'' transit model at a
given period and phase.

@@

\clearpage
\appendix

\section{Mathematical model}\sectlabel{math}

Formally, this can be written for the light curve of the $k$-th star as
\begin{eqnarray}\eqlabel{linear-model}
\bvec{f}_k &=& \bvec{A}\,\bvec{w}_k + \mathrm{noise}
\end{eqnarray}
where
\begin{eqnarray}
\bvec{f}_k &=& \left (\begin{array}{cccc}
    f_{k,1} & f_{k,2} & \cdots & f_{k,N}
\end{array}\right )^\T
\end{eqnarray}
is the list aperture fluxes for star $k$ observed at $N$ times
\begin{eqnarray}
\bvec{t} &=& \left (\begin{array}{cccc}
    t_{1} & t_{2} & \cdots & t_{N}
\end{array}\right )^\T \quad.
\end{eqnarray}
In \eq{linear-model}, the design matrix is given by
\begin{eqnarray}
\bvec{A} &=& \left (\begin{array}{cccccc}
    x_{1,1} & x_{2,1} & \cdots & x_{J,1} & 1 & m_\bvec{\theta}(t_1) \\
    x_{1,2} & x_{2,2} & \cdots & x_{J,2} & 1 & m_\bvec{\theta}(t_2) \\
    && \vdots &&&\\
    x_{1,N} & x_{2,N} & \cdots & x_{J,N} & 1 & m_\bvec{\theta}(t_N)
\end{array}\right )
\end{eqnarray}
where the $x_{j,n}$ are the basis ELCs---with the index $j$ running over
components and the index $n$ running over time---and $m_\bvec{\theta}(t)$ is
the transit model
\begin{eqnarray}
m_\bvec{\theta}(t) &=& \left\{\begin{array}{cl}
-1 & \mathrm{if\,}t\,\mathrm{in\,transit} \\
0 & \mathrm{otherwise}
\end{array}\right.
\end{eqnarray}
parameterized by a period, phase, and transit duration (these parameters are
denoted by \bvec{\theta}).
